---
title: "in-class exercise 1 demo"
author: "xu lin"
date: 2023-11-18
---

##getting start

The code chunk below load the following packages:

-   tmap for thematic mapping

-   sf for geospatial data handling

-   tidyverse for non-spatial data handling

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, dplyr)
```

## Preparing the data

## Importing the 00 data

Firstly, we will import the "*Passenger volume by origin Destination Bus Stops"* data set downlaoded from LTA Datamall by using"read.csv()" of *readr* package

```{r}
obus <- read_csv("data/aspatial/origin_destination_bus_202308.csv")
```

```{r}
obus$ORIGIN_PT_CODE <- as.factor(obus$ORIGIN_PT_CODE)
obus$DESTINATION_PT_CODE <- as.factor(obus$ORIGIN_PT_CODE)
```

```{r}
ortrip_7_9 <- obus %>% filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 7 & TIME_PER_HOUR <=9) %>%
  group_by(ORIGIN_PT_CODE)%>% 
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

## Import bus stop

```{r}
busstop <- st_read(dsn = "data/geospatial", layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(busstop)
```

```{r}
mpsz <- st_read (dsn = "data/geospatial", layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz
```
